
---
title: "Homework 1"
subtitle: "*Evan David (ead955)*"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

We want to prove $(A^{-1})^{T} = (A^T)^{-1}$

Let us define $B = (A^{-1})$

Then $B^T = (A^{-1})^T$

By definition, $AB = A(A^{-1}) = I$, the identity matrix

It follows that $(AB)^T = I^T = I$, or $B^TA^T = I$

Then multiplying both sides by $(A^T)^{-1}$, we have $B^TA^T(A^T)^{-1} = I(A^T)^{-1}$

That gives $B^TI = (A^T)^{-1}$

Then we have $B^T = (A^T)^{-1}$

And so $B^T = (A^{-1})^T = (A^T)^{-1}$

## Problem 2

We will first set up the problem. We'll define $x_1$ as first mortgage loans, $x_2$ as second mortgage loans, $x_3$ as home improvement loans, and $x_4$ as personal overdraft loans.

From the problem we know the following:

1. In total $250 million is lent out. This is represented by the equation:
$$x_1 + x_2 + x_3 + x_4  = 250$$
2. First mortgages are 55% of all mortgages (i.e., first and second mortgage) issued.
$$x_1 = 0.55(x_1 + x_2)$$
Or, $$0.45x_1 - 0.55x_2 = 0$$
3. Second mortgages are 25% of all loans issued. This gives:
$$x_2  = 0.25(x_1 + x_2 + x_3 + x_4)$$
Or, $$-0.25x_1 + 0.75x_2 - 0.25x_3 - 0.25x_4 = 0$$
4. The average interest rate on all loans is 15%. From this, and the information we know about the interest rates of each loan type, we can say:
$$0.14x_1 + 0.2x_2 + 0.2x_3 + 0.1x_4 = 0.15(x_1 + x_2 + x_3 + x_4)$$
Or, $$-0.01x_1 + 0.05x_2 + 0.05x_3 - 0.05x_4 = 0$$
From these equations we can set up a matrix equation $$Ax = y$$

$$\left[\begin{array}
{rrrr}
1 & 1 & 1 & 1 \\
0.45 & -0.55 & 0 & 0 \\
-0.25 & 0.75 & -0.25 & -0.25 \\
-0.01 & 0.05 & 0.05 & -0.05
\end{array}\right]\left[\begin{array}
{r}
x_1 \\
x_2 \\
x_3 \\
x_4
\end{array}\right] = \left[\begin{array}
{r}
250 \\
0 \\
0 \\
0
\end{array}\right]$$
Now we solve for x using matrix inversion:

$$x = A^{-1}y$$

```{r, include=FALSE}
A = matrix(c(1,1,1,1,.45,-.55,0,0,-.25,.75,-.25,-.25,-.01, .05,.05,-.05),nrow = 4,ncol = 4)
A = t(A)
y = matrix(c(250,0,0,0))
x = solve(A) %*% y
```
The solution then gives:
```{r, echo=FALSE}
x
```
And these are the values of $x_1$, $x_2$, $x_3$, and $x_4$, in millions, that give the best lending strategy.

## Problem 3

In this problem, we want to maximize profit by determining the best number of units manufactured for each variant. 

### Decision Variables
$x_1$, $x_2$, $x_3$, and $x_4$ will represent the number of units manufactured for each type of variant.

### Constraints
First, the non-negative constraints:
$$x_1, x_2, x_3, x_4 \geqslant 0$$
Additionally, we have constraints for assembly, polishing, and packing, which are as follows:
$$2x_1 + 4x_2 + 3x_3 + 7x_4 \leqslant 100000$$
$$3x_1 + 2x_2 + 3x_3 + 4x_4 \leqslant 50000$$
$$2x_1 + 3x_2 + 2x_3 + 5x_4 \leqslant 60000$$

### Objective

We want to maximize profit, so we will maximize:
$$1.5x_1 + 2.5x_2 + 3x_3 + 4.5x_4$$

## Problem 4

### Part a

The following R code will generate a 20 by 20 Lehmer matrix A:
```{r}
A = matrix(0,nrow = 20,ncol = 20)

for (i in 1:nrow(A)) {
  for (j in 1:ncol(A)){
    A[i,j] = min(i,j)/max(i,j)
    
  }
}
```
### Part b

Test whether A is symmetric, by checking if $A = A^T$
```{r}
# checks whether the transpose of A is equal to A
all.equal(t(A),A)
```
Therefore A is symmetric.

### Part c

```{r}
C = solve(A)
all.equal(C %*% A, diag(20))
```
Therefore C is the correct inverse of A.

### Part d

```{r}
d = c(1:10,10:1)
d
```
### Part e

$$x = A^{-1}Cd$$
```{r}
x = solve(A) %*% C %*% d
x
```

